Differential invariant signatures for planar Lie group transformations of images

### 2020-10-06 paper/newpaper.tex
Bug fix
Words

% REQUIRED
\begin{keywords}
\end{keywords}

@@

Shapes/objects/images can be acted on by simple transformations that change
their appearance. We want to recognise the objects or identify the set of
transformations. If the set is a group, then we can seek to either identify
the group or solve the equivalence problem by
\begin{enumerate}
    \item Registration, or
    \item Invariance.
\end{enumerate}
The first is commonly studied but is computationally expensive and prone to
local minima. If you're not interested in the actual transformation it is
therefore to be avoided. Here we'll consider invariance. There are many
choices of type e.g. Fourier, integral, or differential invariants. The
last set are cheap to compute and have nice theoretical properties but are
not robust to noise. However, we'll use them anyway. This paper is a
tutorial on differential invariants for planar Lie groups (e.g.
Euclidean, affine, projective) since they are relatively simple, useful,
and relevant to images. We'll use these to demonstrate the methods of
constructing and using differential invariant signatures. Note that this
reduces the recognition problem to recognising higher-dimensional
differential invariant signatures. We'll mention some approaches to this.
See the companion paper for an implementation.

Some examples, plus mention Kendall shape.

The setting that we will consider is a transformation group $\mathcal{G}$
acting on a manifold $\mathcal{M}$ of dimension $n$. A $d$-th order
differential invariant is a local scalar function $I: J^{d} (\mathbb{R},
\mathcal{M} ) \to \mathbb{R}$ that is invariant under the action of
$\mathcal{G}$, i.e., $I(g(x,u^{(d)})) = I(x, u^{(d)}) \forall g \in
\mathcal{G}$, where $J^{d}$ denotes the $d$-th order jet space.

\subsection{Transformation notation}
Most of the time we will work on grey-scale transformations of two-dimensional images. In this setting, we define the transformation group action by a coordinate transformation $\varphi \colon \mathbb{R}^2 \to \mathbb{R}^2$:
\begin{equation}\label{eq:transformation}
  \mathbf{x} \mapsto \bar{\mathbf{x}} = \varphi(\mathbf{x})
\end{equation}
The transformed image is then defined to be:
\begin{equation}\label{eq:fbarequalsf}
  \bar{f}(\bar{\mathbf{x}}) = (f\circ\varphi^{-1})(\bar{\mathbf{x}}) = f(\mathbf{x}).
\end{equation}
Derivatives of $\bar{f}$ and $f$ can be related by the chain rule, for example the first order derivatives are given by:
\begin{equation}\label{eq:prolongation}
  \begin{aligned}
    f_x &= \bar{f}_{\bar{x}} \bar{x}_x + \bar{f}_{\bar{y}}\bar{y}_x \\
    f_y &= \bar{f}_{\bar{x}} \bar{x}_y + \bar{f}_{\bar{y}}\bar{y}_y \\
  \end{aligned}
\end{equation}
where $\bar{x}_x = \frac{\partial \varphi_1}{\partial x}(x, y)$, et cetera. It is often convenient to work with $\varphi^{-1}$ instead, as in \S\ref{sec:movingframes}.

Mention colour here.

\subsection{Stuff from dispaper to tidy}
====
Let $G$ be a (finite or infinite-dimensional) local planar transformation
group.  A {\em $k$-colour image} is a triple $(f,\Omega,k)$ where
$\Omega\subset\R^2$, $1\le k\in\mathbb{Z}$, and $f\colon\Omega\to\R^k$. The
function $f$ will be taken to be as smooth as necessary.  A transformation
$\varphi\in G$ acts on images by $\varphi\cdot (f,\Omega,k) =
(f\circ\varphi^{-1},\varphi(\Omega),k)$.  (As we work locally, we will
usually omit the domain $\Omega$; if the number of colours $k$ is not
given, we assume $k=1$.)  A differential invariant is an invariant of the
prolonged action of $G$ on the derivatives of $f$, i.e., a function $I$ of
the values of $f$ and its derivatives at a point such that $I(f)(x,y) =
I(\varphi\cdot f)(\varphi(x,y))$. An {\em $m$-dimensional signature set} is
the image of $m$ invariants $\mathcal{I}=(I_1,I_2,\dots,I_m)$, i.e. the set
$\mathcal{I}(f)(\Omega)\subset \mathbb{R}^m$.  A signature set is complete
for $f$ if it determines $f$ up to transformations from $G$, i.e. if
$\mathcal{I}(\tilde f)(\tilde\Omega) = \mathcal{I}(f)(\Omega)$ implies
there exists a transformation $\varphi\in G$ such that $\tilde f =
\varphi\cdot f$ and $\tilde \Omega = \varphi(\Omega)$.

Signatures

Idea, some notation, then completeness, bad set, have not removed the recognition problem, just removed the group variation.

---
TEXT FROM OUR MOBIUS PAPER
Calabi et al. [9] propose the use of differential invariant signatures for shape analysis, and further
argue that these should be approximated in a group-invariant way. For example, for the Euclidean
group, the signature is the shape (; s) regarded as a subset of R2. The claimed advantages
of the approach are that the signature determines the shape; that it does not depend on the
choice of initial point on the curve or on parameterization by arclength, and that its G-invariance
makes it robust; and that it is based on a general procedure for arbitrary objects and groups.
See [2, 4, 9, 18, 30, 35, 36] for further developments and applications of the invariant signature.
Although the signature at first sight appears to be complete (e.g., Theorem 5.2 in [9]),
a more detailed treatment (e.g., [17, 28]) highlights the fact that it is not complete on shapes
that contain singular parts { straight and circular segments in the Euclidean case. For these
parts, the signature reduces to a point. Thus, for shapes that have nearly straight or nearly
circular segments, the signature cannot be robust. In addition, while the signature does not
depend on the starting point or the parameterization, it takes values in a very complicated set,
namely, the planar shapes. To compare the invariants of two shapes requires comparing two
shapes. Essentially, the parameterization-dependence has only been deferred to a later stage
of the analysis (unless one is content to compare shapes visually). One approach to this is to
weight the signature, see [32] for more details.
---

@@

@@

local invariants: occlusions OK
derivatives: noise is not

signatures, dimensions, orders
Bad set, completeness

@@

For example, up to second derivatives, we have the following Euclidean invariants, expressed as tensor contractions:
\begin{align*}
  f &= f \\
  f_if_i &= f_x^2 + f_y^2 \\
  f_{ii} &= f_{xx}^2 + f_{yy}^2 \\
  f_{ij} f_{ij} &= f_{xx}^2 + 2f_{xy}^2 + f_{yy}^2 \\
  f_i f_j f_{ij} &= f_x^2 f_{xx} + 2f_x f_y f_{xy} + f_y^2 f_{yy}
\end{align*}

@@

One way to compute invariants is to compare the action of group elements on separate copies of the underlying space. For images, this is typically $\mathbb{R}^2$; colour images in $\mathbb{R}^{2k}$ will be considered in Section~\ref{sec:aaah}.

@@

\begin{equation}
\Omega_{ij} \left(f(x_i, y_i) g(x_j, y_j)\right) = f_{x_i}g_{y_j} - f_{y_i}g_{x_j}
\end{equation}

@@

\begin{equation}
\mbox{tr} \left[\left( \Pi_{k=1}^r \Omega_{i_k j_k} \right) f^{(1)}(x_1, y_1)
f^{(2)}(x_2, y_2), \ldots f^{(n)} (x_n, y_n)\right]
\end{equation}
where $i_k \neq j_k \in \{1, \ldots, n\}$.

@@

For example, consider a weight $2$ transvectant defined as
\begin{align*}
    \mbox{tr} &\left[(\Omega_{12}\Omega_{12}) f(x_1, y_1)f(x_2, y_2)\right] \\
    &= \mbox{tr}\left[(\partial_{x_1 y_2} - \partial_{x_2y_1})(\partial_{x_1 y_2} -
    \partial_{x_2y_1})f(x_1, y_1)f(x_2, y_2)\right]\\
    &= \mbox{tr}\left[\left(\partial_{x_1 y_2} - \partial_{x_2 y_1}\right)\left(f_x(x_1, y_1)f_y(x_2, y_2) - f_y(x_1, y_1)f_x(x_2, y_2)\right)\right] \\
    &= \mbox{tr}\left[f_{xx}(x_1, y_1)f_{yy}(x_2, y_2) - f_{xy}(x_1, y_1)f_{xy}(x_2, y_2)\right] \\
    &= f_{xx}(x, y) f_{yy}(x, y) - f_{xy}(x, y)^2
\end{align*}
This is a degree $2$ partial transvectant as it involves derivatives of $f$
up to second order, and can be represented graphically by the following
diagram,
\begin{center}
\begin{tikzpicture}[baseline=0]
    \node[draw,circle,minimum size=0.2cm] (1) at (-0.5,0) {1};
    \node[draw,circle,minimum size=0.2cm] (2) at (0.5,0) {2};
    \draw[-] (1) to [out=30,in=150] (2);
    \draw[-] (1) to [out=-30,in=-150] (2);
\end{tikzpicture}
\end{center}
where the double edge between nodes $1$ and $2$ corresponds to the two copies of $\Omega_{12}$.
Tables~\ref{tab:transvectants1} and~\ref{tab:transvectants2} shows all possible nonzero partial
transvectants of $A_2$ up to weight $4$ that can be generated in this way.
Note that the weight is the number of edges in the graph.

@@

\todo{Explain}

\subsection{Signatures of $SA(2)$ and $A(2)$}
Recall that invariants of a supergroup are automatically invariants of any of
its subgroups. For example, invariants of $A(2)$ are automatically invariants of $SA(2)$, but not necessarily vice versa. Any of the partial transvectants found in Tables~\ref{tab:transvectants1} and~\ref{tab:transvectants2} are invariants of $SA(2)$ (as the determinant is $1$) but only up to scale factor ($\det A^{-r}$) of $A(2)$.

Hence, we can form a signature for $SA(2)$ by choosing any combination of
partial transvectants. We'll choose a signature with three components so that
it can be visualised as a three-dimensional surface; note that this will not
be complete and may well have a very large bad set. For example, we can choose the function value together with the first two partial transvectants to form the signature:
\begin{equation*}
    \left(f, f_{xx}f_{yy}-f_{xy}^2, f_x^2f_{yy} - 2f_xf_yf_{xy} + f_{xx}f_y^2\right)
\end{equation*}

@@

====

Discussion of singularities.

And then a list of explicit invariants by group.

And the signatures.

@@

\todo[inline]{This needs a friendly introduction, or at least a pointer to the notebooks}

Idea is that it is the transformation back to the cross-section.

The method of moving frames enables one to identify invariants under the action of groups of transformations. It has a long history, as summarised in \cite{Olver14}; particularly noteworthy contributions are by Cartan~\cite{Cartan37} and Olver and co-authors; see~\cite{Olver14} for an overview, and note particularly for an application to~\cite{Calabi98}. The method is described in detail in~\cite{OlverCIT}.

A right moving frame is a smooth $\mathcal{G}$-equivariant map $\rho : \mathcal{M} \to \mathcal{G}$ such that $\rho (g \cdot z) = \rho(z) \cdot g^{-1}$.

In order to construct differential invariants to the action of Lie group $\mathcal{G}$ acting on manifold $\mathcal{M}$, the procedure can be summarised as follows:

\begin{enumerate}
\item Prolong the group action to the jet space of $n$-th order derivatives (using implicit differentiation): $F: J^n \to \mathbb{R}$. 
\item Apply Cartan normalisation:
    \begin{enumerate}
    \item Choose a local cross-section to the group orbits, i.e., a ($n - \dim \mathcal{G}$)-dimensional submanifold $\mathcal{K}$ that intersects transversally at most once with each orbit
    \item $\mathcal{K}$ is specified by $\dim \mathcal{G}$ independent equations $Z_i (z) = c_i$, for $z \in J^n$,  and $Z_i$ scalar-valued functions and $c_i$ constants.
    \item The right moving frame $g = \rho(z)$ is found by solving the normalisation questions $Z_i (g \cdot z) = c_i$ for the group parameters $g$ (in terms of $z$).
    \end{enumerate}
\item The differential invariants are not affected by the action of the group: $I(g \cdot z) = I(z) \: \forall z \in \mathrm{dom} I$, or equivalently, if it is constant on the orbits. 
\item Hence $F$ is made invariant by transforming it with a group action: $I(z) = F(g \cdot z) |_{g=\rho(z)}$.  
\end{enumerate}

Olver uses the group $SE(2)$ on curves as an example~\cite{Olver}, recovering the Euclidean curvature as a second-order invariant. As an example, we use it to compute invariants to $SE(2)$ on the plane, before moving on to two other planar Lie group examples: the M\"obius group ($PSL(2,\mathbb{C})$) and the projective group ($PSL(3,\mathbb{R})$).

\subsection{The Special Euclidean Group $SE(2)$}
An element of the special Euclidean group $SE(2)$ acts on an element $(x, y)$ of $\mathbb{R}^2$ to produce $(\bar{x}, \bar{y})$ as 
\begin{equation*}
  \begin{aligned}
    \bar{x} &= x\cos\theta  - y\sin\theta + t_x \\
    \bar{y} &= x\sin\theta  + y\cos\theta + t_y.
  \end{aligned}
\end{equation*}
We prolong the group action to form the jet space $J^2$ with coordinates $(x,
y, f, f_x, f_y, f_{xx}, f_{xy}, f_{yy})$ via the chain rule. The
transformation rule for the derivatives is given by
\begin{equation}
  \begin{bmatrix}
  \bar f_{\bar x} \\ \bar f_{\bar y} \\ \bar f_{\bar{x}\bar{x}} \\ \bar f_{\bar{x}\bar{y}} \\ \bar f_{\bar{y}\bar{y}}
  \end{bmatrix} = 
  \begin{bmatrix}
 c & -s & 0 & 0 & 0 \\
 s & c & 0 & 0 & 0 \\
0 & 0 & c^2 & -2cs & s^2 \\
0 & 0 & cs & c^2 - s^2 & -cs \\
0 & 0 & s^2 & 2cs & c^2 \\
  \end{bmatrix}
  \begin{bmatrix}
f_x \\ f_y \\ f_{xx} \\ f_{xy} \\ f_{yy} 
  \end{bmatrix},
\label{eqn:SE2}
 \end{equation}
where $c = \cos \theta$ and $s = \sin \theta$. Because the group has three
parameters, we choose the cross section of the group orbits to be $\bar{x}
= 0, \bar{y} = 0, \bar{f}_{\bar{y}} = 0$. We also require that $\bar{f}_{x}
> 0$ to uniquely define the moving frame. Note that the group action is
locally free (away from critical points) when prolonged to $J^1$, however
we prolonged to $J^2$ in order to compute invariants. points. The moving
frame is then given by the prolonged Euclidean transformation that maps an
element of  $J^2$ to this cross section. It can be thought of as rotating
the image about that point so that its gradient is pointing in the positive
$x$ direction, and then translating the image so the point is mapped to the
origin.

@@

The remaining elements not fixed by the cross section, $\bar{f}$,
$\bar{f}_{\bar{x}}$, $\bar{f}_{\bar{x}\bar{x}}$,
$\bar{f}_{\bar{x}\bar{y}}$, and $\bar{f}_{\bar{y}\bar{y}}$ are therefore
all invariant. This gives the following set of invariants:
\begin{equation}\label{eq:se2invariants}
\begin{aligned}
  I_0 &= f \\
  I_1 &= f_x^2 + f_y^2 & &(= \bar{f}_{\bar{x}}^2)\\
  I_2 &= f_x^2 f_{xx} + 2 f_x f_y f_{xy} + f_y^2 f_{yy} &&(= I_1\bar{f}_{\bar{x}\bar{x}}) \\
  I_3 &= f_x f_y (f_{yy} - f_{xx}) + f_{xy} (f_x^2 - f_y^2) &&(= I_1\bar{f}_{\bar{x}\bar{y}})\\
  I_4 &= f_y^2 f_{xx} - 2 f_x f_y f_{xy} + f_x^2 f_{yy} &&(= I_1\bar{f}_{\bar{y}\bar{y}})
\end{aligned}
\end{equation}
Note that we multiplied each of the second derivative invariants by $I_1$
so that they would decay smoothly to zero at critical points.

@@

We again prolong the group action to $J^2$, but this time we expect to need
to use a second derivative term to determine $\varepsilon$ for the moving
frame. The derivative transformations are given by
\begin{equation}
  \begin{bmatrix}
  \bar f_{\bar x} \\ \bar f_{\bar y} \\ \bar f_{\bar{x}\bar{x}} \\ \bar f_{\bar{x}\bar{y}} \\ \bar f_{\bar{y}\bar{y}}
  \end{bmatrix} = 
  \begin{bmatrix}
 \varepsilon c & -\varepsilon s & 0 & 0 & 0 \\
 s & c & 0 & 0 & 0 \\
0 & 0 & c^2 & -2cs & s^2 \\
0 & 0 & \varepsilon cs & \varepsilon(c^2 - s^2) & -\varepsilon cs \\
0 & 0 & s^2 & 2cs & c^2 \\
  \end{bmatrix}
  \begin{bmatrix}
f_x \\ f_y \\ f_{xx} \\ f_{xy} \\ f_{yy} 
  \end{bmatrix},
\label{eqn:E2prolongation}
\end{equation}
where again $c = \cos\theta$ and $s = \sin\theta$.

>>>>

% REQUIRED
\begin{keywords}

\end{keywords}

@@

Biologists differentiate between the {\em form} of an object, and the {\em shape}, being the part of form that is unaffected by size. More generally, the appearance of a fixed object in a set of images can vary markedly depending upon such variables as change of pose and relative position to the camera, even though the object itself does not change. Such variations highlight the importance of `shape', as opposed to `form', in object recognition.

It is natural to seek representations of objects that vary only with regard to shape change, rather than to any change in form. If the space of objects in denoted $M$ and the non-shape transformations form a group $\mathcal{G}$, then mathematically one seeks a way to study objects in $M / \mathcal{G}$. This could be done, for example, by transforming all objects to match a chosen target one by registration. Hence for the group of similarity transformations (rotation and reflection, translation, and scale), a Procrustes alignment~\cite{} will transform all of the shapes into the reference frame of one template example. However, analysis on these quotient spaces is rarely simple~\cite{}. An alternative option is to seek $\mathcal{G}$-invariant functions on $M$ and use the theory of invariants, which is what we consider here.

Invariant signatures -- self-equivalences
Different sorts of invariant, and their properties.
local invariants: occlusions OK
derivatives: noise is not

have not removed the recognition problem, just removed the group variation.

%Shapes/objects/images can be acted on by simple transformations that change
%their appearance. We want to recognise the objects or identify the set of
%transformations. If the set is a group, then we can seek to either identify
%the group or solve the equivalence problem by
%\begin{enumerate}
    %\item Registration, or
    %\item Invariance.
%\end{enumerate}
%The first is commonly studied but is computationally expensive and prone to
%local minima. If you're not interested in the actual transformation it is
%therefore to be avoided. Here we'll consider invariance. There are many
%choices of type e.g. Fourier, integral, or differential invariants. The
%last set are cheap to compute and have nice theoretical properties but are
%not robust to noise. However, we'll use them anyway.

This paper is a
tutorial on differential invariants for planar Lie groups (e.g.
Euclidean, affine, projective) since they are relatively simple, useful,
and relevant to images. We'll use these to demonstrate the methods of
constructing and using differential invariant signatures. Note that this
reduces the recognition problem to recognising higher-dimensional
differential invariant signatures. We'll mention some approaches to this.
See the companion paper for an implementation.

\subsection{Notation}

We define a {\em $k$-colour image} as a triple $(f,\Omega,k)$ where $\Omega\subset\R^n$ (typically, images are planar, so $n=2$), $1\le k\in\mathbb{Z}$ is the number of colour channels, and $f\colon\Omega\to\R^k$. The function $f$ will be taken to be as smooth as necessary. Consider a (finite or infinite-dimensional) local transformation group $\mathcal{G}$.  A transformation $\varphi\in G$ acts on images by $\varphi\cdot (f,\Omega,k) = (f\circ\varphi^{-1},\varphi(\Omega),k)$. For many of our examples we consider greyscale images, so $k=1$, but we discuss colour images in Section~\ref{sec:colour}. Also, since we work locally, we will usually omit the domain $\Omega$.

In order to define differential invariants, it is necessary to identify how the group transformation $\varphi$ acts on derivatives of the function $f$. This is known as the jet space, or jet bundle~\cite{OlverJets}. The $d$-th order jet space $J^d$ consists of the function itself and the set of all partial derivatives of the function up to order $d$. (Note that there are $\left( \begin{array}{c} d+k-1 \\ k \end{array} \right)$ partial derivatives of order $d$). As the group $\mathcal{G}$ is local, it acts naturally on derivatives of the transformation, and so each individual transformation $\varphi \in \mathcal{G}$ can be prolonged to act on the jet space $J^d$. We can thus prolong the action of the group and consider the action of $\mathcal{G}$ on not just $f$, but all derivatives of $f$ up to order $d$. A $d$-th order differential invariant is then a local scalar function $I: J^{d} (f,\Omega, k) \to \mathbb{R}$ that is invariant under the action of $\mathcal{G}$, i.e., $I(\varphi(f,\Omega, k)) = I(f, \Omega, k) \forall \varphi \in \mathcal{G}$.

An $m$-dimensional signature set is the image of $m$ invariants $\mathcal{I}=(I_1,I_2,\dots,I_m)$, i.e., the set $\mathcal{I}(f)(\Omega)\subset \mathbb{R}^m$.  A signature set is complete for $f$ if it determines $f$ up to transformations from $\mathcal{G}$, i.e., if $\mathcal{I}(\tilde f)(\tilde\Omega) = \mathcal{I}(f)(\Omega)$ implies that there exists a transformation $\varphi\in \mathcal{G}$ such that $\tilde f = \varphi\cdot f$ and $\tilde \Omega = \varphi(\Omega)$. Where the signature set is not complete, it is still desirable that the set of objects that are indistinguishable from other be as small as possible. Useful properties for invariants are discussed in detail in~\cite{UsMobius}.

NEED TO TALK ABOUT WHEN IT IS SINGULAR

For the case of two-dimensional greyscale images it is often convenient to work in coordinates. In this case the transformation group action by a coordinate transformation $\varphi \colon \mathbb{R}^2 \to \mathbb{R}^2$ is given by:
\begin{equation}\label{eq:transformation}
  \mathbf{x} \mapsto \bar{\mathbf{x}} = \varphi(\mathbf{x}).
\end{equation}
The transformed image is then defined to be:
\begin{equation}\label{eq:fbarequalsf}
  \bar{f}(\bar{\mathbf{x}}) = (f\circ\varphi^{-1})(\bar{\mathbf{x}}) = f(\mathbf{x}).
\end{equation}
Derivatives of $\bar{f}$ and $f$ can be related by the chain rule, for example the first order derivatives are given by:
\begin{equation}\label{eq:prolongation}
  \begin{aligned}
    f_x &= \bar{f}_{\bar{x}} \bar{x}_x + \bar{f}_{\bar{y}}\bar{y}_x \\
    f_y &= \bar{f}_{\bar{x}} \bar{x}_y + \bar{f}_{\bar{y}}\bar{y}_y \\
  \end{aligned}
\end{equation}
where $\bar{x}_x = \frac{\partial \varphi_1}{\partial x}(x, y)$, et cetera. It is often convenient to work with $\varphi^{-1}$ instead, as in \S\ref{sec:movingframes}.

\subsection{Relevant Literature}
---
TEXT FROM OUR MOBIUS PAPER
Calabi et al. [9] propose the use of differential invariant signatures for shape analysis, and further
argue that these should be approximated in a group-invariant way. For example, for the Euclidean
group, the signature is the shape \todo[inline]{something didn't copy paste correctly here} (??; ??s) regarded as a subset of R2. The claimed advantages
of the approach are that the signature determines the shape; that it does not depend on the
choice of initial point on the curve or on parameterization by arclength, and that its G-invariance
makes it robust; and that it is based on a general procedure for arbitrary objects and groups.
See [2, 4, 9, 18, 30, 35, 36] for further developments and applications of the invariant signature.
Although the signature at first sight appears to be complete (e.g., Theorem 5.2 in [9]),
a more detailed treatment (e.g., [17, 28]) highlights the fact that it is not complete on shapes
that contain singular parts { straight and circular segments in the Euclidean case. For these
parts, the signature reduces to a point. Thus, for shapes that have nearly straight or nearly
circular segments, the signature cannot be robust. In addition, while the signature does not
depend on the starting point or the parameterization, it takes values in a very complicated set,
namely, the planar shapes. To compare the invariants of two shapes requires comparing two
shapes. Essentially, the parameterization-dependence has only been deferred to a later stage
of the analysis (unless one is content to compare shapes visually). One approach to this is to
weight the signature, see [32] for more details.
---

@@

JUSTIFY THIS SET OF GROUPS

@@

@@

For example, up to second derivatives, we have the following Euclidean invariants, expressed as tensor contractions:
\begin{align*}
  f &= f \\
  f_if_i &= f_x^2 + f_y^2 \\
  f_{ii} &= f_{xx}^2 + f_{yy}^2 \\
  f_{ij} f_{ij} &= f_{xx}^2 + 2f_{xy}^2 + f_{yy}^2 \\
  f_i f_j f_{ij} &= f_x^2 f_{xx} + 2f_x f_y f_{xy} + f_y^2 f_{yy}
\label{eq:se2}
\end{align*}

@@

Another way to compute invariants is to compare the action of group elements on separate copies of the underlying space (here, $\mathbb{R}^2$).

@@

\begin{equation}
\Omega_{ij} \left(f(x_i, y_i) g(x_j, y_j)\right) = f_{x_i}g_{y_j} - f_{y_i}g_{x_j}.
\end{equation}

@@

\begin{equation}
\mbox{tr} \left[\left( \Pi_{k=1}^r \Omega_{i_k j_k} \right) f^{(1)}(x_1, y_1)
f^{(2)}(x_2, y_2), \ldots f^{(n)} (x_n, y_n)\right],
\end{equation}
where $i_k \neq j_k \in \{1, \ldots, n\}$.

@@

For example, consider a weight $2$ transvectant defined as:
\begin{align*}
    \mbox{tr} &\left[(\Omega_{12}\Omega_{12}) f(x_1, y_1)f(x_2, y_2)\right] \\
    &= \mbox{tr}\left[(\partial_{x_1 y_2} - \partial_{x_2y_1})(\partial_{x_1 y_2} -
    \partial_{x_2y_1})f(x_1, y_1)f(x_2, y_2)\right]\\
    &= \mbox{tr}\left[\left(\partial_{x_1 y_2} - \partial_{x_2 y_1}\right)\left(f_x(x_1, y_1)f_y(x_2, y_2) - f_y(x_1, y_1)f_x(x_2, y_2)\right)\right] \\
    &= \mbox{tr}\left[f_{xx}(x_1, y_1)f_{yy}(x_2, y_2) - f_{xy}(x_1, y_1)f_{xy}(x_2, y_2)\right] \\
    &= f_{xx}(x, y) f_{yy}(x, y) - f_{xy}(x, y)^2.
\end{align*}
This is a degree $2$ partial transvectant as it involves derivatives of $f$
up to second order, and can be represented graphically by the following
diagram:
\begin{center}
\begin{tikzpicture}[baseline=0]
    \node[draw,circle,minimum size=0.2cm] (1) at (-0.5,0) {1};
    \node[draw,circle,minimum size=0.2cm] (2) at (0.5,0) {2};
    \draw[-] (1) to [out=30,in=150] (2);
    \draw[-] (1) to [out=-30,in=-150] (2);
\end{tikzpicture}
\end{center}
where the double edge between nodes $1$ and $2$ corresponds to the two copies of $\Omega_{12}$.
Tables~\ref{tab:transvectants1} and~\ref{tab:transvectants2} shows all possible nonzero partial
transvectants of $A_2$ up to weight $4$ that can be generated in this way.
Note that the weight is the number of edges in the graph.

@@

\todo[inline]{Explain}

\subsection{Signatures of $SA(2)$ and $A(2)$}
Recall that invariants of a supergroup are automatically invariants of any of
its subgroups. For example, invariants of $A(2)$ are automatically invariants of $SA(2)$, but not necessarily vice versa. Any of the partial transvectants found in Tables~\ref{tab:transvectants1} and~\ref{tab:transvectants2} are invariants of $SA(2)$ (as the determinant is $1$), but only up to scale factor ($\det A^{-r}$) of $A(2)$.

Hence, we can form a signature for $SA(2)$ by choosing any combination of
partial transvectants. We'll choose a signature with three components so that
it can be visualised as a three-dimensional surface; note that this will not
be complete and may well have a very large bad set. For example, we can choose the function value together with the first two partial transvectants to form the signature:
\begin{equation*}
    \left(f, f_{xx}f_{yy}-f_{xy}^2, f_x^2f_{yy} - 2f_xf_yf_{xy} + f_{xx}f_y^2\right).
\end{equation*}

@@

@@

So far the methods that we have seen to can construct differential invariants for particular groups are not directly generalisable. It is natural to ask if there is an algorithmic approach that will yield invariants more directly. The answer is a partial yes. There is a general method, based on Cartan's method of moving frames, that is generally applicable, but finding good solutions with it still requires some ingenuity in general. We introduce the method here and then demonstrate its use to find invariants to further groups.

A moving frame is a mapping into a group $\mathcal{G}$ that provides a frame of reference along a curve. Cartan's method of moving frames generalises the concept to a homogeneous space of a Lie group $\mathcal{G}$. When defined on the principal bundle it defines a $\mathcal{G}$-equivariant mapping, which enables one to identify invariants under the action of groups of transformations. The method of moving frames has a long history, as summarised in \cite{Olver14}; particularly noteworthy contributions are by Cartan~\cite{Cartan37} and Olver and co-authors; see~\cite{Olver14} for an overview, and note particularly for an application to~\cite{Calabi98}. The method is described in detail in~\cite{OlverCIT}.

A (right-) moving frame is a smooth $\mathcal{G}$-equivariant map $\rho : \mathcal{M} \to \mathcal{G}$ such that $\rho (g \cdot z) = \rho(z) \cdot g^{-1}$, where $z \in J^d$. The procedure to construct differential invariants to the action of Lie group $\mathcal{G}$ acting on manifold $\mathcal{M}$ using the method of moving frames is as follows:

\begin{enumerate}
\item Prolong the group action to the jet space of $d$-th order derivatives (using implicit differentiation): $F: J^d \to \mathbb{R}$. 
\item Apply Cartan normalisation:
    \begin{enumerate}
    \item Choose a local cross-section to the group orbits, i.e., a ($d - \dim \mathcal{G}$)-dimensional submanifold $\mathcal{K}$ that intersects transversally at most once with each orbit
    \item $\mathcal{K}$ is specified by $\dim \mathcal{G}$ independent equations $Z_i (z) = c_i$, for $z \in J^d$,  and $Z_i$ scalar-valued functions and $c_i$ constants.
    \item The right moving frame $g = \rho(z)$ is found by solving the normalisation questions $Z_i (g \cdot z) = c_i$ for the group parameters $g$ (in terms of $z$). In other words, the moving frame is the transformation back to the cross-section.
    \end{enumerate}
\item The differential invariants are not affected by the action of the group: $I(g \cdot z) = I(z) \: \forall z \in \mathrm{dom} I$, or equivalently, if it is constant on the orbits. 
\item Hence $F$ is made invariant by transforming it with a group action: $I(z) = F(g \cdot z) |_{g=\rho(z)}$.  
\end{enumerate}

Olver uses the group $SE(2)$ on curves as an example~\cite{Olver}, recovering the Euclidean curvature as a second-order invariant. We continue this example to compute invariants to $SE(2)$ on the plane, before moving on to other planar Lie group examples: $E(2)$, the M\"obius group ($PSL(2,\mathbb{C})$) and the projective group ($PSL(3,\mathbb{R})$). Finally, we apply it to an infinite-dimensional pseudo-Lie group, the conformal diffeomorphisms.

\subsection{The Special Euclidean Group $SE(2)$}

An element of the special Euclidean group $SE(2)$ acts on an element $(x, y)$ of $\mathbb{R}^2$ to produce $(\bar{x}, \bar{y})$ as:
\begin{equation*}
  \begin{aligned}
    \bar{x} &= x\cos\theta  - y\sin\theta + t_x \\
    \bar{y} &= x\sin\theta  + y\cos\theta + t_y.
  \end{aligned}
\end{equation*}
We prolong the group action to form the jet space $J^2$ with coordinates $(x,
y, f, f_x, f_y, f_{xx}, f_{xy}, f_{yy})$ via the chain rule. The
transformation rule for the derivatives is given by:
\begin{equation}
  \begin{bmatrix}
  \bar f_{\bar x} \\ \bar f_{\bar y} \\ \bar f_{\bar{x}\bar{x}} \\ \bar f_{\bar{x}\bar{y}} \\ \bar f_{\bar{y}\bar{y}}
  \end{bmatrix} = 
  \begin{bmatrix}
 c & -s & 0 & 0 & 0 \\
 s & c & 0 & 0 & 0 \\
0 & 0 & c^2 & -2cs & s^2 \\
0 & 0 & cs & c^2 - s^2 & -cs \\
0 & 0 & s^2 & 2cs & c^2 \\
  \end{bmatrix}
  \begin{bmatrix}
f_x \\ f_y \\ f_{xx} \\ f_{xy} \\ f_{yy} 
  \end{bmatrix},
\label{eqn:SE2}
 \end{equation}
where $c = \cos \theta$ and $s = \sin \theta$. Because the group has three
parameters, we choose the cross section of the group orbits to be $\bar{x}
= 0, \bar{y} = 0, \bar{f}_{\bar{y}} = 0$. We also require that $\bar{f}_{x}
> 0$ to uniquely define the moving frame. Note that the group action is
locally free (away from critical points) when prolonged to $J^1$, however
we prolonged to $J^2$ in order to compute invariants. points. The moving
frame is then given by the prolonged Euclidean transformation that maps an
element of  $J^2$ to this cross section. It can be thought of as rotating
the image about that point so that its gradient is pointing in the positive
$x$ direction, and then translating the image so the point is mapped to the
origin.

@@

The remaining elements not fixed by the cross section, $\bar{f}$,
$\bar{f}_{\bar{x}}$, $\bar{f}_{\bar{x}\bar{x}}$,
$\bar{f}_{\bar{x}\bar{y}}$, and $\bar{f}_{\bar{y}\bar{y}}$ are therefore
all invariant. This gives the following set of invariants:
\begin{equation}\label{eq:se2invariants}
\begin{aligned}
  I_0 &= f \\
  I_1 &= f_x^2 + f_y^2 & &(= \bar{f}_{\bar{x}}^2)\\
  I_2 &= f_x^2 f_{xx} + 2 f_x f_y f_{xy} + f_y^2 f_{yy} &&(= I_1\bar{f}_{\bar{x}\bar{x}}) \\
  I_3 &= f_x f_y (f_{yy} - f_{xx}) + f_{xy} (f_x^2 - f_y^2) &&(= I_1\bar{f}_{\bar{x}\bar{y}})\\
  I_4 &= f_y^2 f_{xx} - 2 f_x f_y f_{xy} + f_x^2 f_{yy} &&(= I_1\bar{f}_{\bar{y}\bar{y}}).
\end{aligned}
\end{equation}
Note that we multiplied each of the second derivative invariants by $I_1$
so that they would decay smoothly to zero at critical points.

@@

We again prolong the group action to $J^2$, but this time we expect to need
to use a second derivative term to determine $\varepsilon$ for the moving
frame. The derivative transformations are given by:
\begin{equation}
  \begin{bmatrix}
  \bar f_{\bar x} \\ \bar f_{\bar y} \\ \bar f_{\bar{x}\bar{x}} \\ \bar f_{\bar{x}\bar{y}} \\ \bar f_{\bar{y}\bar{y}}
  \end{bmatrix} = 
  \begin{bmatrix}
 \varepsilon c & -\varepsilon s & 0 & 0 & 0 \\
 s & c & 0 & 0 & 0 \\
0 & 0 & c^2 & -2cs & s^2 \\
0 & 0 & \varepsilon cs & \varepsilon(c^2 - s^2) & -\varepsilon cs \\
0 & 0 & s^2 & 2cs & c^2 \\
  \end{bmatrix}
  \begin{bmatrix}
f_x \\ f_y \\ f_{xx} \\ f_{xy} \\ f_{yy} 
  \end{bmatrix},
\label{eqn:E2prolongation}
\end{equation}
where again $c = \cos\theta$ and $s = \sin\theta$.
